---
title: "Shiny_app"
author: "Mihir-Riya"
date: "2025-11-28"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This R Markdown document is made interactive using Shiny. Unlike the more traditional workflow of creating static reports, you can now create documents that allow your readers to change the assumptions underlying your analysis and see the results immediately. 

To learn more, see [Interactive Documents](http://rmarkdown.rstudio.com/authoring_shiny.html).

## Inputs and Outputs

You can embed Shiny inputs and outputs in your document. Outputs are automatically updated whenever inputs change.  This demonstrates how a standard R plot can be made interactive by wrapping it in the Shiny `renderPlot` function. The `selectInput` and `sliderInput` functions create the input widgets used to drive the plot.

```{r eruptions, echo=FALSE}
inputPanel(
  selectInput("n_breaks", label = "Number of bins:",
              choices = c(10, 20, 35, 50), selected = 20),
  
  sliderInput("bw_adjust", label = "Bandwidth adjustment:",
              min = 0.2, max = 2, value = 1, step = 0.2)
)

renderPlot({
  hist(faithful$eruptions, probability = TRUE, breaks = as.numeric(input$n_breaks),
       xlab = "Duration (minutes)", main = "Geyser eruption duration")
  
  dens <- density(faithful$eruptions, adjust = input$bw_adjust)
  lines(dens, col = "blue")
})
```

## Embedded Application

It's also possible to embed an entire Shiny application within an R Markdown document using the `shinyAppDir` function. This example embeds a Shiny application located in another directory:

```{r tabsets, echo=FALSE}
shinyAppDir(
  system.file("examples/06_tabsets", package = "shiny"),
  options = list(
    width = "100%", height = 550
  )
)
```

Note the use of the `height` parameter to determine how much vertical space the embedded application should occupy.

You can also use the `shinyApp` function to define an application inline rather then in an external directory.

In all of R code chunks above the `echo = FALSE` attribute is used. This is to prevent the R code within the chunk from rendering in the document alongside the Shiny components.


```{r}
if (!require("janitor")) install.packages("janitor")
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("lubridate")) install.packages("lubridate")
if (!require("shinydashboard")) install.packages("shinydashboard")
if (!require("shiny")) install.packages("shiny")
if (!require("reshape2")) install.packages("reshape2")
if (!require("DT")) install.packages("DT")
if (!require("plotly")) install.packages("plotly")
if (!require("visNetwork")) install.packages("visNetwork")

library(tidyverse)
library(janitor)
library(lubridate)
library(shiny)
library(shinydashboard)
library(reshape2)
library(DT)
library(plotly)
library(visNetwork)
```

```{r}
df <- read.csv("DIG.csv") %>%
  janitor::clean_names()
df

df <- df %>%
  select(id, trtmt, age, race, sex, ejf_per, chestx, bmi, klevel, creat, digdoser, elevjvp, pedema, restdys, exertdys, actlimit, pulcong, heartrte, diabp, sysbp, prevmi, angina, diabetes, hyperten, diguse, diuretk, diuret, ksupp, aceinhib, nitrates, hydral, vasod, digdose, cvd, whf, mi, uang, strk, sva, vena, crev, ocvd, rinf, oth, hosp, hospdays, nhosp, death, deathday) %>%
  
  mutate(
    id = as.character(id),
    trtmt = as.factor(trtmt),
    age = as.numeric(age),
    race = as.factor(race),
    sex = as.factor(sex),
    ejf_per = as.numeric(ejf_per),
    chestx = as.numeric(chestx),
    bmi = as.numeric(bmi),
    klevel = as.numeric(klevel),
    creat = as.numeric(creat),
    digdoser = as.numeric(digdoser),
    elevjvp = as.factor(elevjvp),
    pedema = as.factor(pedema),
    restdys = as.factor(restdys),
    exertdys = as.factor(exertdys),
    actlimit = as.factor(actlimit),
    pulcong = as.factor(pulcong),
    heartrte = as.numeric(heartrte),
    diabp = as.numeric(diabp),
    sysbp = as.numeric(sysbp),
    prevmi = as.factor(prevmi),
    angina = as.factor(angina),
    diabetes = as.factor(diabetes),
    hyperten = as.factor(hyperten),
    diguse = as.factor(diguse),
    diuretk = as.factor(diuretk),
    diuret = as.factor(diuret),
    ksupp = as.factor(ksupp),
    aceinhib = as.factor(aceinhib),
    nitrates = as.factor(nitrates),
    hydral = as.factor(hydral),
    vasod = as.factor(vasod),
    digdose = as.numeric(digdose),
    cvd = as.factor(cvd),
    whf = as.factor(whf),
    mi = as.factor(mi),
    uang = as.factor(uang),
    strk = as.factor(strk),
    sva = as.factor(sva),
    vena = as.factor(vena),
    crev = as.factor(crev),
    ocvd = as.factor(ocvd),
    rinf = as.factor(rinf),
    oth = as.factor(oth),
    hosp = as.factor(hosp),
    hospdays = as.numeric(hospdays),
    nhosp = as.numeric(nhosp),
    death = as.factor(death),
    deathday = as.numeric(deathday))
df


df <- df %>%
  mutate(
    Treatment = factor(trtmt, levels = c(0, 1), labels = c("Placebo", "Digoxin")),
    Race = factor(race, levels = c(1, 2), labels = c("White", "Non-white")),
    Sex = factor(sex, levels = c(1, 2), labels = c("Male", "Female")),
    ElevatedJVP = factor(elevjvp, levels = c(0, 1, 2, 3), labels = c("Normal", "Mild", "Moderate", "Severe")),
    Pedema = factor(pedema, levels = c(0, 1, 2, 3), labels = c("Normal", "Mild", "Moderate", "Severe")),
    Dyspnea_at_rest = factor(restdys, levels = c(0, 1, 2, 3), labels = c("Normal", "Mild", "Moderate", "Severe")),
    Dyspnea_on_exertion = factor(exertdys, levels = c(0, 1, 2, 3), labels = c("Normal", "Mild", "Moderate", "Severe")),
    Action_limitation = factor(actlimit, levels = c(0, 1, 2, 3), labels = c("Normal", "Mild", "Moderate", "Severe")),
    Pulmonary_congestion = factor(pulcong, levels = c(0, 1, 2, 3), labels = c("Normal", "Mild", "Moderate", "Severe")),
    Prior_MI = factor(prevmi, levels = c(0, 1), labels = c("No", "Yes")),
    Angina = factor(angina, levels = c(0, 1), labels = c("No", "Yes")),
    Diabetes = factor(diabetes, levels = c(0, 1), labels = c("No", "Yes")),
    HTN = factor(hyperten, levels = c(0, 1), labels = c("No", "Yes")),
    On_Digoxin = factor(diguse, levels = c(0, 1), labels = c("No", "Yes")),
    On_Ksparing = factor(diuretk, levels = c(0, 1), labels = c("No", "Yes")),
    On_Diuretic = factor(diuret, levels = c(0, 1), labels = c("No", "Yes")),
    On_Ksupp = factor(ksupp, levels = c(0, 1), labels = c("No", "Yes")),
    On_ACE = factor(aceinhib, levels = c(0, 1), labels = c("No", "Yes")),
    On_Nitrates = factor(nitrates, levels = c(0, 1), labels = c("No", "Yes")),
    On_Hydralazine = factor(hydral, levels = c(0, 1), labels = c("No", "Yes")),
    On_Vasodilator = factor(vasod, levels = c(0, 1), labels = c("No", "Yes")),
    CVD = factor(cvd, levels = c(0, 1), labels = c("No", "Yes")),
    WHF = factor(whf, levels = c(0, 1), labels = c("No", "Yes")),
    MI = factor(mi, levels = c(0, 1), labels = c("No", "Yes")),
    Unstable_Angina = factor(uang, levels = c(0, 1), labels = c("No", "Yes")),
    Stroke = factor(strk, levels = c(0, 1), labels = c("No", "Yes")),
    Supraventricular_Arrhythmia = factor(sva, levels = c(0, 1), labels = c("No", "Yes")),
    Ventricular_Arrythmia = factor(vena, levels = c(0, 1), labels = c("No", "Yes")),
    Coronary_Revasc = factor(crev, levels = c(0, 1), labels = c("No", "Yes")),
    Other_CVD = factor(ocvd, levels = c(0, 1), labels = c("No", "Yes")),
    Respiratory_infection = factor(rinf, levels = c(0, 1), labels = c("No", "Yes")),
    Other_noncardiac = factor(oth, levels = c(0, 1), labels = c("No", "Yes")),
    Hospitalization = factor(hosp, levels = c(0, 1), labels = c("No", "Yes")),
    Death = factor(death, levels = c(0, 1), labels = c("Alive", "Dead"))) %>%
  rename(
    ID = id,
    Age = age,
    Prescribed_dose = digdose,
    Recommended_Dig_dose = digdoser, 
    BMI = bmi,
    EF_percent = ejf_per,
    ChestXR_ratio = chestx,
    Heart_rate = heartrte,
    Diastolic_BP = diabp,
    Systolic_BP = sysbp,
    Hospital_days = hospdays,
    No.of_Hosp = nhosp,
    Death_or_Hosp_WHF = deathday,
    )

# ------------------------------------------------------------------------------------------------------------------
# starting with the shiny app development

library(shinydashboard)


# this is created for bar plot of symptoms (specific subset)
df <- df %>%
  mutate(
    ElevatedJVP = factor(elevjvp, levels = c(0, 1, 2, 3), labels = c("Normal", "Mild", "Moderate", "Severe")),
    Pedema = factor(pedema, levels = c(0, 1, 2, 3), labels = c("Normal", "Mild", "Moderate", "Severe")),
    Dyspnea_at_rest = factor(restdys, levels = c(0, 1, 2, 3), labels = c("Normal", "Mild", "Moderate", "Severe")),
    Dyspnea_on_exertion = factor(exertdys, levels = c(0, 1, 2, 3), labels = c("Normal", "Mild", "Moderate", "Severe")),
    Action_limitation = factor(actlimit, levels = c(0, 1, 2, 3), labels = c("Normal", "Mild", "Moderate", "Severe")),
    Pulmonary_congestion = factor(pulcong, levels = c(0, 1, 2, 3), labels = c("Normal", "Mild", "Moderate", "Severe"))
  )



#-------------------------------------------------------------------------------------------------------------------
# Adding ui to the shiny app
ui <- dashboardPage(
  dashboardHeader(title = "Digitalis Randomized, Double-blind, Multicenter Trial", titleWidth = 550), # trial title is given to the shiny dashboard header
  
  
#Two side bar is created main and explore
  dashboardSidebar(
    sidebarMenu(
      menuItem("Main", tabName = "Main", icon = icon("dashboard")),
      menuItem("Explore", tabName = "Explore", icon = icon("chart-bar"))),
    
    fileInput("file1", "Choose a File"), verbatimTextOutput("file1_contents"), # File upload function added to the side bar
    
    
    selectInput(inputId = "Treatment", label = "Select Participants Treatment", choices = c("Placebo", "Digoxin"), multiple = TRUE),
    
    radioButtons(inputId = "Sex", label = "Participant's Gender", choices = c("Male", "Female")),
    
    sliderInput(inputId = "Age", "Participant's Age:", min = 18, max = 95, value = c(20, 30)),
    
    checkboxGroupInput("checkGroup", label = h3("Cardiovascular Symptoms"), choices = list("ElevatedJVP" = 1, "Pedema" = 2, "Dyspnea_at_rest" = 3, "Dyspnea_on_exertion"= 4, "Action_limitation" = 5, "Pulmonary_congestion" = 6)),
    
    hr(), fluidRow(column(3, verbatimTextOutput("value"))),
    
    downloadButton("downloadData", "Download DIG")), #download button for dig data
#----------------------------------------------------------------------------------------------------------------------
# Dashboard --> Main page, added 2 summary table  
  dashboardBody(
    tabItems(
      tabItem(tabName = "Main",
               fluidRow(
                
                 box(width = 12, title = "Trial Data with Participant's Demographics", collapsible = TRUE, status = "warning", solidHeader = TRUE, dataTableOutput("table1")),
                 
                 box(width = 12, title = "Cardiovascular conditions of Participants", collapsible = TRUE, status = "warning", solidHeader = TRUE, dataTableOutput("table2"))
      )
    ),
#---------------------------------------------------------------------------------------------------------------------
# Dashboard --> Explore page, added multiple plots and a table
    tabItem(tabName = "Explore",
            fluidRow(
              box(width = 6, title = "Summary", collapsible = TRUE, solidHeader = TRUE, status = "primary", tableOutput("SummaryTable")),
              
              
              
              box(width = 6, title = "Interative Heatmap", collapsible = TRUE, solidHeader = TRUE, status = "primary", plotlyOutput("corrplot", height = "550px")),
              
              
              
              box(width = 5, title = "Box Plot", collapsible = TRUE, status = "warning", solidHeader = TRUE, plotlyOutput("plot1")),
              
              
              
              box(width = 5, title = "Cardiovascular Symptoms", collapsible = TRUE, status = "primary", solidHeader = TRUE, plotlyOutput("barPlot")),
              
              
              box(width = 6, title = "CVD-Drug Network", collapsible = TRUE, status = "primary", solidHeader = TRUE, visNetworkOutput("CVDnetwork", height = "500px")),
              
              
              box(width = 5, title = "Violin Plot", collapsible = TRUE, status = "primary", solidHeader = TRUE, plotlyOutput("violinPlot"))
              )
            )
    ) # closing tabitems parentheses
  ) #closing dashboardbody parentheses
) #closing dashboardPage parentheses
    
#---------------------------------------------------------------------------------------------------------------------
# Building server for shiny app
server <- function(input, output) {

  # filtering reactive data
  df_sub <- reactive({
    df %>%
      filter(Treatment %in% input$Treatment) %>%
      filter(Sex == input$Sex) %>%
      filter(Age>= input$Age[1] & Age <= input$Age[2])
    })

  output$value <- renderPrint({input$checkGroup})

#--------------------------------------------------------------------------------    
# File input
 
  output$file1_contents <- renderPrint({print(input$file1)})

  
   
#----------------------------------------------------------------------------------  
# Download button
  
  output$downloadData <- downloadHandler(
    filename = "DIG.csv",
    content = function(file) {
      write.csv(DIG, file)
    })

# Data Table  1
  output$table1 <- renderDataTable({
    df_sub() %>%
      select(
        ID, Sex, Age, Race, BMI, Treatment, HTN, Diabetes, CVD, On_Digoxin, Prescribed_dose, Recommended_Dig_dose
      )
  })

# Data Table  2
  output$table2 <- renderDataTable({
    df_sub() %>%
      select(
        ID, EF_percent, klevel, ElevatedJVP, Pedema, Dyspnea_at_rest, HTN, Diabetes, CVD, Prior_MI, Angina, Stroke, Death
      )
  })  
  
#---------------------------------------------------------------------------------------------------------  
#Summary Table  
  output$SummaryTable <- renderTable({
    st <- df_sub()
    req(nrow(st) > 0)
    
    num_df <- st[, sapply(st, is.numeric)]
    req(ncol(num_df) > 0)
      
    data.frame(
      Variable = names(num_df),
      Mean = sapply(num_df, mean, na.rm = TRUE),
      Median = sapply(num_df, median, na.rm = TRUE),
      SD = sapply(num_df, sd, na.rm = TRUE),
      Min = sapply(num_df, min, na.rm = TRUE),
      Max = sapply(num_df, max, na.rm = TRUE))
    
  })
 
#--------------------------------------------------------------------------------------------------------- 
# violin plot
  
  output$violinPlot <- renderPlotly({
    st <- df_sub() %>%
      select(Sex, klevel, creat, EF_percent, ChestXR_ratio, Heart_rate, Diastolic_BP, Systolic_BP)
    
    st_vio <- st %>%
      pivot_longer(
        cols = -Sex,
        names_to = "Variable",
        values_to = "Value")
    
    fig <- plot_ly(
      data = st_vio,
      x = ~Variable,
      y = ~Value,
      color = ~Sex,
      type = "violin",
      split = ~Sex,
      box = list(visible = TRUE),
      meanline = list(visible = TRUE)
      )
    
    fig <- fig %>%
      layout(
        yaxis = list(zeroline = FALSE),
        violinmode = "group"
      )
}) 
      
  
#--------------------------------------------------------------------------------------------
 # bar plot 
output$barPlot <- renderPlotly({
  req(input$checkGroup)

  vars_map <- c(
    "1" = "ElevatedJVP",
    "2" = "Pedema",
    "3" = "Dyspnea_at_rest",
    "4" = "Dyspnea_on_exertion",
    "5" = "Action_limitation",
    "6" = "Pulmonary_congestion"
  )

  selected_vars <- vars_map[input$checkGroup]

  st <- df_sub() %>% select(all_of(selected_vars))
  req(nrow(st) > 0)

  # Convert to long format
  long_df <- tidyr::pivot_longer(st,
                                 cols = everything(),
                                 names_to = "Symptom",
                                 values_to = "Severity")

  # Plot
  plot_ly(
    data = long_df,
    x = ~Symptom,
    color = ~Severity,
    type = "bar"
  ) %>%
    layout(
      barmode = "stack",
      title = "Severity Distribution of CV Symptoms",
      xaxis = list(title = "Symptom"),
      yaxis = list(title = "Count of Participants")
    )
})

  
 #------------------------------------------------------------------------------------------------------------ 
# Boxplot  
  output$plot1 <- renderPlotly({
    st <- df_sub() %>%
      select(CVD, Prescribed_dose)
    
    req(nrow(st) > 0)
    
    plot_ly(data = st,
            x = ~CVD, y = ~Prescribed_dose, type = "bar", color = ~as.factor(Prescribed_dose)) %>%
      layout(barmode = "stack",
             title = "Prescribed dose for CVD patients",
             xaxis = list(title = "CVD Status"),
             yaxis = list(title = "Count"),
             legend = list(title = list(text = "Prescribed Dose")))
    
  })

  
    
#--------------------------------------------------------------------------------------------------
# heatmap (explore)  
  output$corrplot <- renderPlotly({
    st <- df_sub()
    req(nrow(st) > 1) # rows should be more than one
    
    num_df <- st[, sapply(st, is.numeric)] #only numeric columns
    
    req(ncol(num_df) > 1) #need at least 2 numeric vars
    
    corr <- cor(num_df, use = "pairwise.complete.obs")
    
    plot_ly(
      x = colnames(corr),
      y = rownames(corr),
      z = corr,
      type = "heatmap",
      colorscale = list(
        c(0,1),
        c("blue", "red")),
      zmin = -1,
      zmax = 1) %>%
      layout(
        xaxis = list(tickangle = 45),
        yaxis = list(autorange = "reversed")
      )
  })
  
  
#-----------------------------------------------------------------------------------------------
#CVD network plot 
  
  output$CVDnetwork <- renderVisNetwork({
    st <- df_sub() %>%
      select(CVD, On_Digoxin, On_Ksparing, On_Diuretic, On_Ksupp, On_ACE, On_Nitrates, On_Hydralazine, On_Vasodilator)
    
    req(nrow(st) > 0)
    
    drug_cols <- c("On_Digoxin", "On_Ksparing", "On_Diuretic", "On_Ksupp", "On_ACE", "On_Nitrates", "On_Hydralazine", "On_Vasodilator")
    
    
    edges <- lapply(drug_cols, function(drug) {
      st %>%
        group_by(CVD) %>%
        summarise(count = sum(.data[[drug]] == "Yes")) %>%
        mutate(drug = drug)
    }) %>% bind_rows()
    
    edges <- edges %>% filter(count > 0) 
    
# network edges created    
    edges_df <- edges %>% 
      mutate(from = CVD, to = drug, value = count, title = paste0(count, "patients")) %>%
      select(from, to, value, title)

# node list created    
    nodes_df <- data.frame(
      id = c("Yes", "No", drug_cols),
      label = c("CVD", "Non-CVD", drug_cols),
      color = c("red", "green", rep("lightblue", length(drug_cols))),
      shape = "ellipse"
    )


# Intreactive graph
    visNetwork(nodes_df, edges_df) %>%
      visEdges(smooth = TRUE) %>%
      visLayout(randomSeed = 123) %>%
      visOptions(highlightNearest = TRUE, nodesIdSelection = TRUE)

  })
}
shinyApp(ui = ui, server = server)
```
